---
title: "Mini-Project 2"
author: "Amy Park & Kate Bruncati"
date: "2019-03-24"
output: 
  html_document:
    code_folding: show
---


## Loading the data

```{r, message=FALSE}
library(tidyverse)
library(fs)
install.packages("wordcountaddin")
wordcountaddin::word_count("mp2.Rmd")
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
glimpse(contributions)
```

Contributions are the most common form of campaign support. A contribution is anything of value given, loaned, or advanced to influence a federal election. The data utilized in our project originated from the individual contribution or contributions set based on the 2011-2012 election cycles. The information given represented contributions from an individual to a federal committee/candidate. 

As seen in the past, a candidate is received much more favorably by the public if he/she has enough resources - funds - to strengthen his/her advertisements and overall campaign strategy. Equipped with the proper budget to finance certain propaganda, it is nearly inevitable that the candidate with the highest amount of contributions will have the most ‘pull’ with the public and ultimately, are more likely to be elected.  

```{r}
committees_transaction_amt <- full_join(candidates, contributions, by = "cand_id") %>%
  select(cmte_id, transaction_amt, cand_name, cand_election_yr, cand_party_affiliation) %>% 
  filter(transaction_amt > 0) %>%
  group_by(cand_name, cand_election_yr, cand_party_affiliation) %>%
  summarize(total_trans_amt = sum(transaction_amt)) %>%
  arrange(desc(total_trans_amt))
head(committees_transaction_amt, 20)
transaction_20 <- head(committees_transaction_amt, 20)
```

To prove this, our first set of data presented - via a bar chart - is meant to represent the 20 candidates who have received the most in terms of contributions. Because candidates’ transaction amounts differ in various parts of the year, we chose to take a sum of the amounts for each candidate in order to compare, overall, which candidate had the largest contributions. 

```{r}
transaction_20$cand_party_affiliation <- as.factor(transaction_20$cand_party_affiliation)
graph_1 <-
  ggplot(transaction_20,
         aes(x = reorder(cand_name, total_trans_amt), 
             y = total_trans_amt)) + 
  geom_bar(stat = "identity", aes(fill = factor(cand_party_affiliation))) +
  geom_text(aes(label = total_trans_amt, hjust = -.1)) + 
  coord_flip() + 
  expand_limits(y = 500000000) + 
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual("Party Affiliation", values = c("red","blue")) +
  labs(x = NULL, y = NULL, 
       title = "Top 20 Contributions in 2012", 
       subtitle = "Highest Contribution among Candidates in 2012")
show(graph_1)
```

Our data revealed Barack Obama as the person with the highest transaction amount for the 2012 election year. Such a result fully supports the information presented beforehand  that candidates are more likely to win as their contributions grow. 8 Republican and 7 Democratic candidates are easily depicted in their signature blue and red party colors. 

#Second Part of Project

In the second part of our project, we chose to focus on primary and general votes. The two elections form a two-step process and are used to determine the winner. The first of these is the primary election, which serves as a party nomination for the political office. Each political party is allowed only one candidate in the presidential election. Therefore, the primary election is the means to determine the party’s one official presidential candidate. After the primary elections, a general election will be held. This is the main election which is held between the winners of the various primaries. It will typically result in a much higher voter turnout than the primary elections. 

```{r}
house_elections %>%
  filter(general_votes > 0, primary_votes > 0, party == "R" | party == "D") %>%
  select(party, primary_votes, general_votes, ge_winner) %>%
  filter(!is.na("ge_winner")) %>%
  ggplot(aes(x = primary_votes, y = general_votes)) +
  facet_wrap(~party) + geom_point(stat = "identity", aes(color = factor(ge_winner))) + geom_smooth() + 
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Primary Votes", y = "General Votes", 
       title = "Relationship between Primary Votes and General Votes", 
       subtitle = "House Elections in the U.S.") + 
  scale_color_manual("Win?", values = c("goldenrod1", "dodgerblue1"))
```

This part of the data represents the relationship between primary and general votes for Republicans and Democrats. Some candidates are shown to have zero primary or general votes for certain reasons; for example, a candidate with zero primary votes might be the only candidate for that party. Our data have been set to only include votes larger than zero. 

The second graph - a scatterplot -  reveals that the relationship between general and primary votes is directly proportional with a strong correlation between the two and, as the number of votes increase, they disperse further. The winning candidates are color-coded as blue and the losing candidates as yellow, respectively. While there is a clear correlation between the number of votes a candidate receives, it appears to be weaker among the Democratic Party, which is an interesting observation we took note of.

> Word count: `r wordcountaddin::word_count()`